---
title: "Test Performance for Bivariate Cells"
author: "Aneesha Iqbal"
date: "09/06/2023"
output: html_document
---

# Install Packages

```{r setup, include=FALSE}
library(spatstat)
library(rearrr)
```


# Function to generate m ellipse cells containing independent bivariate 
# homogeneous poisson process
```{r}
m_indep_bv_hpp <- function(m, l){
  cells <- c()
  for (i in 1:m){
    a <- runif(1)
    b <- runif(1)
    # adjusted intensity = l/(a*b*pi), events marked "i" and "j"
    bv_pp <- rmpoispp(l/(a*b*pi), types=c("i","j"), win=ellipse(a,b))
    xy_c <- coords(bv_pp)
    # mapping ellipses to circles
    xy_c$x <- xy_c$x/a
    xy_c$y <- xy_c$y/b
    xy_c$labs <- marks(bv_pp)
    cells <- c(cells, xy_c)
  }
  return(cells)
}
```

# Functions used to generate bivariate neyman-scott cells
```{r}
# code taken from help on "rPoissonCluster"
# multitype Neyman-Scott process (each cluster is a multitype process)
nclust2 <- function(x0, y0, radius, n, types=c("i", "j")) {
    X <- runifdisc(n, radius, centre=c(x0, y0))
    M <- sample(types, n, replace=TRUE)
    marks(X) <- M
    return(X)
}

MT_NS_process <- function(kappa, expand, radius, n){
  a <- runif(1) # x-scale
  b <- runif(1) # y-scale
  NS_pp <- rPoissonCluster(kappa/(pi*a*b), expand*sqrt(pi*a*b), nclust2, radius*sqrt(pi*a*b), n, win=ellipse(a,b))
  pp_in_c <- affine(NS_pp, mat=diag(c(1/a, 1/b)))
  xy_c <- coords(pp_in_c)
  xy_c$labs <- marks(pp_in_c)
  return(xy_c)
}

m_MT_NS_process <- function(m, kappa, expand, radius, n){
  all_cells <- c()
  for(j in 1:m){
    xy_df <- MT_NS_process(kappa, expand, radius, n)
    all_cells <- c(all_cells, xy_df)
  }
  return(all_cells)
}
```

# Bivariate point to nearest event distribution (BPNED)
```{r}
PNED_bv <- function(set_of_cells){
  PNED_df <- data.frame("i"=c(0), "j"=c(0))
  m <- length(set_of_cells)/3
  for (i in 1:m){
    labs_i <- set_of_cells[[3*i]]
    xy_c <- data.frame(x=set_of_cells[[3*i-2]], y=set_of_cells[[3*i-1]])

    k <- round(sqrt(nrow(xy_c)))
    xg <- seq(-1, 1, length.out = k)
    yg <- seq(-1, 1, length.out = k)
    xy_grid <- expand.grid(xg, yg)
    k_grid <- xy_grid[which(xy_grid[,1]^2+xy_grid[,2]^2<1),]
    unif_pp <- ppp(k_grid[,1], k_grid[,2], window=disc(1))

    xy1 <- xy_c[which(labs_i=="i"),]
    xy2 <- xy_c[which(labs_i=="j"),]
    p1 <- ppp(xy1[,1], xy1[,2], window=disc(1))
    p2 <- ppp(xy2[,1], xy2[,2], window=disc(1))

    mid_df <- nncross(unif_pp, p1)
    mid_df[,2] <- nncross(unif_pp, p2)[,1]
    colnames(mid_df) <- c("i", "j")
    PNED_df <- rbind(PNED_df, mid_df)
  }
  PNED_df = PNED_df[-1,]
  return(PNED_df)
}
```

# Implementing test from section 4.2 and 2.3
```{r}
 pned_bv_test <- function(setA, setB){ 
  pned_vals_A <- PNED_bv(setA)
  pned_vals_B <- PNED_bv(setB)
  m <- nrow(pned_vals_A)
  n <- nrow(pned_vals_B)
  N <- n + m
  pooled_pned <- rbind(pned_vals_A, pned_vals_B)
  pooled_mean <- colMeans(pooled_pned)
  pooled_pned$i <- pooled_pned$i - pooled_mean[1]
  pooled_pned$j <- pooled_pned$j - pooled_mean[2]
  pooled_pned <- angle(pooled_pned, x_col="i", y_col="j", origin = c(0,0), origin_col_name=NULL, degrees_col_name="angles")
  pooled_pned$set_label <- c(rep('a', m), rep('b', n))
  sorted_pooled_pned <- pooled_pned[order(pooled_pned$angles),]
  sorted_pooled_pned$angle_rank <- seq(N)
  sorted_pooled_pned$u <- cos(2*pi*sorted_pooled_pned$angle_rank/N)
  sorted_pooled_pned$v <- sin(2*pi*sorted_pooled_pned$angle_rank/N)
  new_set_A <- sorted_pooled_pned[which(sorted_pooled_pned$set_label == "a"), ]
  U <- ((sum(new_set_A$u))^2 + (sum(new_set_A$v))^2) * 2 * (N - 1)/(m*n)
  d <- 1 + (N*(N+1)-6*m*n)/(N*(m-1)*(n-1))
  # included f-distribution since suggested in paper - but worse asymptotic properties so not used
  p_value1 <- 1 - pf(U, d, ((N-3)*d)) 
  p_value2 <- 1 - pchisq(U, df = 2)
  return(c(p_value1, p_value2))
}
```


# level of test for independent bivariate homogeneous poisson process
```{r}
p_vals_hpp <- matrix(0, nrow=2, ncol=100)
for(r in 1:100){
  setA <- m_indep_bv_hpp(20,25)
  setB <- m_indep_bv_hpp(20,25)
  p_vals_hpp[, r] <- pned_bv_test(setA, setB)
}
rownames(p_vals_hpp) <- c("f-dist", "chisq-dist")

rowMeans(p_vals_hpp < 0.05)[2]
# output: chisq-dist:  0.05
```

# level of test for bivariate neyman-scott
```{r}
# level of test (type-I error)
p_vals_hpp <- matrix(0, nrow=2, ncol=100)
for(r in 1:100){
  setA <- m_MT_NS_process(10, 10, 1, 0.01, 3)
  setB <- m_MT_NS_process(10, 10, 1, 0.01, 3)
  p_vals_hpp[, r] <- pned_bv_test(setA, setB)
}
rownames(p_vals_hpp) <- c("f-dist", "chisq-dist")

rowMeans(p_vals_hpp < 0.05)[2]
# output: chisq-dist:  0.09
```

# power of test
```{r}
p_vals_hpp_v_NS <- matrix(0, nrow=2, ncol=100)
for(r in 1:100){
  setA <- m_indep_bv_hpp(20,24)
  setB <- m_MT_NS_process(20, 12, 0.1, 0.05, 4)
  p_vals_hpp_v_NS[, r] <- pned_bv_test(setA, setB)
}
rownames(p_vals_hpp_v_NS) <- c("f-dist", "chisq-dist")[2]
# output: chisq-dist:  0.97
```






