---
title: "Visualising Point Patterns and Cells"
author: "Aneesha Iqbal"
date: "09/06/2023"
output: html_document
---

# Install Packages

```{r setup, include=FALSE}
library(spatstat)
library(ggplot2)
```


# Homogeneous Poisson Process (HPP): Intensity 100
## Generate the HPP
```{r}
h_pp <- rpoispp(100)
```

## Visulaise the HPP - Figure 2.1(a)
```{r}
xy1 <- coords(h_pp)
plot(xy1, pch=19, xlab="",xaxt="n", ylab="",yaxt="n", col="blue")
```


# Inhomogeneous Poisson Process (IHPP): lambda(x,y)=100exp[-4y]
## Generate the IHPP
```{r}
ih_pp <- rpoispp(function(x,y) {100*exp(-4*y)}, 100)
```

## Visualise the IHPP - Figure2.1(b)
```{r}
xy2 <- coords(ih_pp)
plot(xy2, pch=19, xlim=c(0, 1), ylim=c(0, 1), xlab="", xaxt="n", 
     ylab="",yaxt="n", col="blue")
```



# Thomas Processes: kapp=20, mu=10
## Generate 3 thomas processes with different sigmas
```{r}
T1 <- rThomas(20, 0.02, 10) # sigma = 0.02
T2 <- rThomas(20, 0.05, 10) # sigma = 0.05
T3 <- rThomas(20, 0.10, 10) # sigma = 0.1
```

## Visualise T1 - Figure 2.2(a)
```{r}
txy1 <- coords(T1)
plot(txy1, pch=20, xlab="",xaxt="n", ylab="",yaxt="n", col="blue")
```

# Visualise T2 - Figure 2.2(b)
```{r}
txy2 <- coords(T2)
plot(txy2, pch=20, xlab="",xaxt="n", ylab="",yaxt="n", col="blue")
```

# Visualise T3 - Figure 2.2(c)
```{r}
txy3 <- coords(T3)
plot(txy3, pch=20, xlab="",xaxt="n", ylab="",yaxt="n", col="blue")
```



# Matérn I Process kappa = 200
## Generate 2 Matérn I Processes for different hardcore distances
```{r}
set.seed(592)

mpp1 <- rMaternI(200, 0.050) # r_H = 0.050
mpp2 <- rMaternI(200, 0.025) # r_H = 0.025

# plot the first Matérn I Process - Figure 2.3(a)
plot(coords(mpp1), col="blue", pch=16, asp=1, 
     xlab="",xaxt="n", ylab="",yaxt="n",
     xlim=c(0,1), ylim=c(0,1))

# plot second Matérn I Process - Figure 2.3(b)
plot(coords(mpp2), col="blue", pch=16, asp=1, 
     xlab="",xaxt="n", ylab="",yaxt="n",
     xlim=c(0,1), ylim=c(0,1))
```


# Ellipse Cells
## HPPs
### function to generate HPP ellipse cells
```{r}
plot_pp_cells <- function(n, l){
  for (i in 1:n) {
    a <- runif(1) # x-scale 
    b <- runif(1) # y-scale
    
    # generate a HPP in the ellipse
    hpp_i <- rpoispp(l/(a*b*pi), win = ellipse(a,b)) #lambda'= l/(ab pi)
    cell_df_i <- coords(hpp_i) # extract coordinates
    
    # print out plot of cells (use ggplot)
    g_ellipse <- ggplot(cell_df_i, aes(x=x, y=y))
    print(g_ellipse + geom_point(col="blue", size=2) + xlim(-1, 1) + ylim(-1, 1) + annotate("path",
    x=a*cos(seq(0,2*pi,length.out=100)),
    y=b*sin(seq(0,2*pi,length.out=100)))+ theme_void())
    print(nrow(cell_df_i)) # print number of events
    }
}
```

### plot 3 HPP ellipse cells with E[N]=50 - Figure 3.1
```{r}
set.seed(111)
plot_pp_cells(3,50)
```

## Thomas Processes
### function to generate thomas process ellipse cells
```{r}
plot_thom_cells <- function(n, k, scale, mu){
  for (i in 1:n) {
    a <- runif(1) # x-scale
    b <- runif(1) # y-scale
    
    # generate thomas process in ellipse
    ## kappa' = k/(ab pi)
    ## sigma' = sqrt(ab pi)*scale
    thom_i <- rThomas(k/(a*b*pi), sqrt(a*b*pi)*scale, mu, win = ellipse(a,b))
    cell_df_i <- coords(thom_i) # extract coordinates
    
    # print out plots of cells (use ggplot)
    g_ellipse <- ggplot(cell_df_i, aes(x=x, y=y))
    print(g_ellipse + geom_point(col="blue", size=2) + xlim(-1, 1) + ylim(-1, 1) + annotate("path",
    x=a*cos(seq(0,2*pi,length.out=100)),
    y=b*sin(seq(0,2*pi,length.out=100)))+ theme_void())
    print(nrow(cell_df_i)) # print number of events
    }
}
```

### plot 3 HPP ellipse cells with kappa=10, sigma=0.03, mu=5 - Figure 3.2
```{r}
set.seed(112)
plot_thom_cells(3, 10, 0.03, 5)
```


# Irregular Cells
## Set boundary of cells
```{r}
# radius function of an irregular cell
r_mag <- function(thetas, u, t){
    r1 <- u[1]*sin(thetas + t[1]) 
    r2 <- u[2]*sin(2*thetas + t[2]) 
    r3 <- u[3]*sin(3*thetas + t[3]) 
    return(r1 + r2 + r3 + 3)
}

# cell 1 boundary
u <- rev(sort(runif(3)))
t <- runif(3)
thetas <- seq(0,2*pi,length.out=1000)
r <- r_mag(thetas, u, t)
x <- r * cos(thetas)
y <- r * sin(thetas)
# set window of cell 2
WC <- owin(poly = list(x = x, y = y))
# area of cell 1
area_cell <- (0.5*sum(u^2)+9)*pi

# cell 2 boundary
u2 <- rev(sort(runif(3)))
t2 <- runif(3)
r2 <- r_mag(thetas, u2, t2)
x2 <- r2 * cos(thetas)
y2 <- r2 * sin(thetas)
# set window of cell 2
WC2 <- owin(poly = list(x = x2, y = y2))
# area of cell 2
area_cell2 <- (0.5*sum(u2^2)+9)*pi
```

```{r}
# Plot a Thomas Process in cell 1 - Figure 3.3(b)
pp_s <- rThomas(10/(area_cell), sqrt(area_cell)*0.03, 5, win = WC)
plot(x,y, pch=20, cex=0.5, xlab="",xaxt="n", ylab="",yaxt="n")
points(coords(pp_s), pch = 16, cex=1.25, col="blue")

# Plot a HPP in cell 2 - Figure 3.3(a)
pp_s2 <- rpoispp(50/area_cell2, win = WC2)
plot(x2,y2, pch=20, cex=0.5, xlab="",xaxt="n", ylab="",yaxt="n")
points(coords(pp_s2), pch = 16, cex=1.25, col="blue")
```


# Ellipse Cell Mapped to Reference Cell
```{r}
ab <- runif(2)
hpp_E <- rpoispp(50/(ab[1]*ab[2]*pi), win=ellipse(ab[1],ab[2]))
hpp_C <- affine(hpp_E, mat=diag(c(1/ab[1],1/ab[2])))
```

## Visualising ellipse cell before and after mapping to circles
```{r}
# Figure 3.4(a)
g_hpp_ellipse <- ggplot(coords(hpp_E), aes(x=x, y=y))
g_hpp_ellipse + geom_point(col="blue", size=2) + xlim(-1, 1) + ylim(-1, 1) + annotate("path", x=ab[1]*cos(seq(0,2*pi,length.out=100)),y=ab[2]*sin(seq(0,2*pi,length.out=100)))+ theme_bw()
# Figure 3.4(b)
g_hpp_circle <- ggplot(coords(hpp_C), aes(x=x, y=y))
g_hpp_circle + geom_point(col="blue", size=2) + xlim(-1, 1) + ylim(-1, 1) + annotate("path", x=cos(seq(0,2*pi,length.out=100)),y=sin(seq(0,2*pi,length.out=100)))+ theme_bw()
```



# Irregular Cell Mapped to Reference Cell
```{r}
in_cell_to_c <- function(x, y, u, t){
        # takes xy-coordinates of events
        # u and t parameters to set boundary
        
        # initialize empty lists
        all_thetas <- c() # to store arguments of events 
        ordered_x <- c() # to store positive then negative values of x
        ordered_y <- c() # store y corresponding to positive then negative values of x
        
        x1 <- x[which(x>=0)] # positive x-coordinates
        ordered_x <- c(ordered_x, x1) # append positive x-coordinates
        y1 <- y[which(x>=0)] # y corresponding to positive x-coordinates
        ordered_y <- c(ordered_y, y1) # append

        # find arguments of coordiantes with positive x-coordinates
        thetas <- atan(y1/x1)
        all_thetas <- c(all_thetas, thetas)

        x1 <- x[which(x<0)] # negative x-coordinates
        ordered_x <- c(ordered_x, x1) # append negative x-coordinates
        y1 <- y[which(x<0)] # y-coordinates corresponding to negative x
        ordered_y <- c(ordered_y, y1)

        # find arguments of coordiantes with negative x-coordinates
        thetas <- pi + atan(y1/x1)
        all_thetas <- c(all_thetas, thetas)

        r1 <- sqrt(ordered_x^2 + ordered_y^2) # find magnitudes of each event

        r2 <- r_mag(all_thetas, u, t) # find radius of cell boundary

        R <- r2[which(r1 < r2)] # magnitudes of events within cell boundary
        # mapping to circle
        x <- ordered_x[which(r1 < r2)] / R
        y <- ordered_y[which(r1 < r2)] / R
        return(data.frame(x,y))
}
```

```{r}
set.seed(304)
# generate cell boundary
u <- rev(sort(runif(3)))
t <- runif(3)
thetas <- seq(0,2*pi,length.out=1000)
r <- r_mag(thetas, u, t)
x1 <- r * cos(thetas)
y1 <- r * sin(thetas)
WC <- owin(poly = list(x = x1, y = y1))
# area of cell
area_cell <- (0.5*sum(u^2)+9)*pi
# generate HPP in the cell boundary
pp_s <- rpoispp(50/area_cell, win = WC)
xy_c <- in_cell_to_c(pp_s$x, pp_s$y, u, t) # map irregular cell to circle
pp_in_c <- ppp(xy_c[,1], xy_c[,2], window=disc(1))
```

## Visualising irregular cell before and after mapping to reference
```{r}
# Figure 3.5(a)
g_irr <- ggplot(coords(pp_s), aes(x=x, y=y))
g_irr + geom_point(col="blue", size=2) + xlim(min(x1,y1), max(x1,y1)) + ylim(min(x1,y1), max(x1,y1)) + annotate("path", x=x1,y=y1)+ theme_bw()
# Figure 3.5(b)
g_irr_circle <- ggplot(coords(pp_in_c), aes(x=x, y=y))
g_irr_circle + geom_point(col="blue", size=2) + xlim(-1, 1) + ylim(-1, 1) + annotate("path", x=cos(seq(0,2*pi,length.out=100)),y=sin(seq(0,2*pi,length.out=100)))+ theme_bw()
```


# Bivariate Independent Posisson Process in an Ellipse
## Figure 4.1(a)
```{r}
set.seed(304)
a1 <- runif(1) # x-scale
b1 <- runif(1) # y-scale
# generate multitype HPP with labels "i" and "j", 
# lambda'= 25/(a1 * b1 * pi) for both
bv_pp <- rmpoispp(25/(a1*b1*pi), types=c("i","j"), win=ellipse(a1,b1))
pp_in_c <- affine(bv_pp, mat=diag(c(1/a1,1/b1))) # mapped to circle

# define colours of points to plot
point_labs <- rep("red", npoints(bv_pp)) 
point_labs[which(bv_pp$marks == "j")] <- "blue"

# visualise cell
plot(coords(bv_pp), col=point_labs, xlim=c(-1, 1), ylim=c(-1,1), pch=16, cex=1.2, asp=1, xlab="",xaxt="n", ylab="",yaxt="n",axes=F)
points(a1*cos(seq(0,2*pi, length.out = 1000)), b1*sin(seq(0,2*pi, length.out = 1000)), pch=20, cex=0.2)
```

## Figure 4.1(b)
```{r}
set.seed(814)
a2 <- runif(1) # x-scale
b2 <- runif(1) # y-scale
# generate multitype HPP with labels "i" and "j", 
# lambda'= 25/(a2 * b2 * pi) for both
bv_pp2 <- rmpoispp(25/(a2 * b2 * pi), types=c("i","j"), win=ellipse(a2,b2))
pp_in_c2 <- affine(bv_pp2, mat=diag(c(1/a2,1/b2)))

# define colours of points to plot
point_labs2 <- rep("red", npoints(bv_pp2))
point_labs2[which(bv_pp2$marks == "j")] <- "blue"

# visualise cell
plot(coords(bv_pp2), col=point_labs2, xlim=c(-1, 1), ylim=c(-1,1), pch=16, cex=1.2, asp=1, xlab="",xaxt="n", ylab="",yaxt="n",axes=F)
points(a2*cos(seq(0,2*pi, length.out = 1000)), b2*sin(seq(0,2*pi, length.out = 1000)), pch=20, cex=0.2)
```

## Figure 4.1(c)
```{r}
set.seed(613)
a3 <- runif(1) # x-scale
b3 <- runif(1) # y-scale
# generate multitype HPP with labels "i" and "j", 
# lambda'= 25/(a3 * b3 * pi) for both
bv_pp3 <- rmpoispp(25/(a3*b3*pi), types=c("i","j"), win=ellipse(a3,b3))

# define colours of points to plot
point_labs3 <- rep("red", npoints(bv_pp3))
point_labs3[which(bv_pp3$marks == "j")] <- "blue"

# visualise cell
plot(coords(bv_pp3), col=point_labs3, xlim=c(-1, 1), ylim=c(-1,1), pch=16, cex=1.2, asp=1, xlab="",xaxt="n", ylab="",yaxt="n",axes=F)
points(a3*cos(seq(0,2*pi, length.out = 1000)), b3*sin(seq(0,2*pi, length.out = 1000)), pch=20, cex=0.2)
```

# Bivariate Neyman-Scott in an Ellipses

```{r}
# code taken from help on "rPoissonCluster"
# multitype Neyman-Scott process (each cluster is a multitype process)
nclust2 <- function(x0, y0, radius, n, types=c("i", "j")) {
    X <- runifdisc(n, radius, centre=c(x0, y0))
    M <- sample(types, n, replace=TRUE)
    marks(X) <- M
    return(X)
  }
```

```{r}
set.seed(412)

a_vals <- runif(3) # x-scales
b_vals <- runif(3) # y-scales

max_lim <- max(a_vals, b_vals)

MNS1 <- rPoissonCluster(15, 0.1, nclust2, radius=0.075, n=10, win=ellipse(a_vals[1], b_vals[1]))
MNS2 <- rPoissonCluster(15, 0.1, nclust2, radius=0.075, n=10, win=ellipse(a_vals[2], b_vals[2]))
MNS3 <- rPoissonCluster(15, 0.1, nclust2, radius=0.075, n=10, win=ellipse(a_vals[3], b_vals[3]))

cols1 <- rep("red", npoints(MNS1))
cols1[which(MNS1$marks == "j")] <- "blue"

cols2 <- rep("red", npoints(MNS2))
cols2[which(MNS2$marks == "j")] <- "blue"

cols3 <- rep("red", npoints(MNS3))
cols3[which(MNS3$marks == "j")] <- "blue"

# Figure 4.2(a)
plot(coords(MNS1), col=cols1, xlim=c(-max_lim, max_lim), ylim=c(-max_lim, max_lim), pch=16, xlab="",xaxt="n", ylab="",yaxt="n",axes=F)
points(a_vals[1]*cos(seq(0, 2*pi, length.out = 1000)), b_vals[1]*sin(seq(0, 2*pi, length.out = 1000)), pch=20, cex=0.2)
legend("bottomleft", inset=.05, c("i", "j"), col=c("red", "blue"), pch=16)

# extra plot not included
plot(coords(MNS2), col=cols2, xlim=c(-max_lim, max_lim), ylim=c(-max_lim, max_lim), pch=16, xlab="",xaxt="n", ylab="",yaxt="n",axes=F)
points(a_vals[2]*cos(seq(0, 2*pi, length.out = 1000)), b_vals[2]*sin(seq(0, 2*pi, length.out = 1000)), pch=20, cex=0.2)
legend("bottomleft", inset=.05, c("i", "j"), col=c("red", "blue"), pch=16)

# Figure 4.3(b)
plot(coords(MNS3), col=cols3, xlim=c(-max_lim, max_lim), ylim=c(-max_lim, max_lim), pch=16, xlab="",xaxt="n", ylab="",yaxt="n",axes=F)
points(a_vals[3]*cos(seq(0, 2*pi, length.out = 1000)), b_vals[3]*sin(seq(0, 2*pi, length.out = 1000)), pch=20, cex=0.2)
legend("bottomleft", inset=.05, c("i", "j"), col=c("red", "blue"), pch=16)
```











