---
title: "Test performance on Irregular Cells"
author: "Aneesha Iqbal"
date: "05/06/2023"
output: html_document
---

```{r global_options, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width = 5, 
                      fig.height = 3)
if (!require(pacman)){
  install.packages("pacman", repos = "http://cran.us.r-project.org")}
pacman::p_load(knitr, tinytex, stats, tidyverse, forecast, 
               ggplot2, pracma, reshape2, tinytex, spatstat)
```

Generating Boundary of Cell
```{r}
r_mag <- function(thetas, u, t){
    r1 <- u[1]*sin(thetas + t[1]) 
    r2 <- u[2]*sin(2*thetas + t[2]) 
    r3 <- u[3]*sin(3*thetas + t[3]) 
    return(r1 + r2 + r3 + 3)
}
```

Find Circumference of Cell
```{r}
find_circumference <- function(thetas){
  r1 <- u[1]*sin(thetas + t[1]) + u[2]*sin(2*thetas + t[2]) + u[3]*sin(3*thetas + t[3]) + 3
  r2 <- u[1]*cos(thetas + t[1]) + 2*u[2]*cos(2*thetas + t[2]) + 3*u[3]*cos(3*thetas + t[3])
  r <- sqrt(r1^2 + r2^2)
  return(r)
}

circumference <- function(u, t){
  c <- integrate(find_circumference, 0, 2*pi)[[1]]
  return(c)
}
```



Function which takes xy-coordinates of points that lie within irregular cell and maps them to within a unit circle
```{r}
in_cell_to_c <- function(x, y, u, t){
  all_thetas <- c()
  ordered_x <- c()
  ordered_y <- c()

  x1 <- x[which(x>=0)]
  ordered_x <- c(ordered_x, x1)
  y1 <- y[which(x>=0)]
  ordered_y <- c(ordered_y, y1)

  thetas <- atan(y1/x1)
  all_thetas <- c(all_thetas, thetas)

  x1 <- x[which(x<0)]
  ordered_x <- c(ordered_x, x1)
  y1 <- y[which(x<0)]
  ordered_y <- c(ordered_y, y1)

  thetas <- pi + atan(y1/x1)
  all_thetas <- c(all_thetas, thetas)

  r1 <- sqrt(ordered_x^2 + ordered_y^2)

  r2 <- r_mag(all_thetas, u, t)

  R <- r2[which(r1 < r2)]
  x <- ordered_x[which(r1 < r2)] / R
  y <- ordered_y[which(r1 < r2)] / R
  return(data.frame(x,y))
}
```

Function that finds the shortest distance from a point in df1 to a point in df2. Then the PNED function for events inside the unit circle
```{r}
dist_between_sets <- function(df1, df2){
  n <- nrow(df1)
  m <- nrow(df2)
  
  x1 <- df1[,1]
  y1 <- df1[,2]
  
  x2 <- df2[,1]
  x2 <- as.data.frame(lapply(x2, rep, n))
  y2 <- df2[,2]
  y2 <- as.data.frame(lapply(y2, rep, n))
  
  distances <- sqrt((x2-x1)^2 + (y2-y1)^2)
  distances <- apply(distances, 1, min)
  return(distances)
}
#---------
PNED_c <- function(n_cells){
  n <- length(n_cells)/2
  all_ds <- c()
  for (i in 1:n) {
    xy_c <- data.frame(n_cells[[2*i - 1]], n_cells[[2*i]])
    
    k <- round(sqrt(nrow(xy_c)))

    x <- seq(-1, 1, length.out = k)
    y <- seq(-1, 1, length.out = k)

    xy_grid <- expand.grid(x, y)

    xs <- xy_grid[,1]
    ys <- xy_grid[,2]
    
    xk <- xs[which(xs^2 + ys^2<1)]
    yk <- ys[which(xs^2 + ys^2<1)]

    k_grid <- data.frame(xk, yk)

    d <- dist_between_sets(k_grid, xy_c)
    all_ds <- c(all_ds, d)
  }
  return(all_ds)
}
```

# Power of Test as a function of Irregularity
```{r}
set.seed(304)
circumferences <- rep(0,10)
cell_r_vars <- rep(0,10)
areas <- rep(0,10)
powers <- rep(0,10)
for (i in 1:10){
    u <- rev(sort(runif(3)))
    t <- runif(3)
    thetas <- seq(0,2*pi,length.out=1000)
    r <- r_mag(thetas, u, t)
    x <- r * cos(thetas)
    y <- r * sin(thetas)
    WC <- owin(poly = list(x = x, y = y))
    area_cell <- (0.5*sum(u^2)+9)*pi
    areas[i] <- area_cell
    circumferences[i] <- circumference(u,t)
    cell_r_vars[i] <- var(r)
    outcomes <- rep(0,50)
    for (j in 1:50){
        set_A <- c()
        for (p in 1:20){
            pp_s <- rpoispp(50/(area_cell), win = WC)
            x1 <- pp_s[["x"]]
            y1 <- pp_s[["y"]]
            df_xy <- in_cell_to_c(x1, y1, u, t)
            set_A <- c(set_A, df_xy)
        }
        set_B <- c()
        for (p in 1:20){
            pp_s <- rThomas(10/(area_cell), sqrt(area_cell)*0.09, 5, win = WC)
            x1 <- pp_s[["x"]]
            y1 <- pp_s[["y"]]
            df_xy <- in_cell_to_c(x1, y1, u, t)
            set_B <- c(set_B, df_xy)
        }
        pned_A <- PNED_c(set_A)
        pned_B <- PNED_c(set_B)
        outcomes[j] <- ks.test(pned_A, pned_B)$p.value < 0.05
    }
    powers[i] <- mean(outcomes)
}
```

```{r}
circumferences/areas
# output: 0.666428570550276 0.670771988490272 0.664519556288033 
# 0.673669893962963 0.669526804805917 0.6770217381394280.664337604112697 
# 0.6676695723661250.662885773649998 0.659851552543466

cell_r_vars
# output: 0.721472449269 0.273733576538683 0.0741522302780681 0.639308920016978 
# 0.814457834530209 0.744882850782814 0.111233957343338 0.670155751020617 
# 0.299722274317176 0.356381155943414

powers
# output: 0.94 0.98 1 0.92 0.94 0.94 0.98 0.96 1 0.94
```

```{r}
# Figure 3.8(a)
plot(cell_r_vars, powers, xlab="variance in radius", 
     ylab="power", main="Power for Cells with different Radius Variances", pch=16)

# Figure 3.8(b)
plot(circumferences/areas, powers, xlab="circumference-to-area ratio", 
     ylab="power", main="Power for Cells with different Circumference-to-Area Ratios", pch=16)
```


# Level of Test

## HPP
```{r}
Thetas <- seq(0,2*pi,length.out=1000)
L_outcomes <- rep(0,100)
for (j in 1:100){
    set_A <- c()
    for (p in 1:20){
        u <- rev(sort(runif(3)))
        t <- runif(3)
        r <- r_mag(Thetas, u, t)
        x <- r * cos(Thetas)
        y <- r * sin(Thetas)
        cell_area <- (0.5*sum(u^2)+9)*pi
        pp_s <- rpoispp(50/(cell_area), win=c(min(x),max(x),min(y),max(y)))
        x1 <- pp_s[["x"]]
        y1 <- pp_s[["y"]]
        df_xy <- in_cell_to_c(x1, y1, u, t)
        set_A <- c(set_A, df_xy)
    }
    set_B <- c()
    for (p in 1:20){
        u <- rev(sort(runif(3)))
        t <- runif(3)
        r <- r_mag(Thetas, u, t)
        x <- r * cos(Thetas)
        y <- r * sin(Thetas)
        cell_area <- (0.5*sum(u^2)+9)*pi
        pp_s <- rpoispp(50/(cell_area), win=c(min(x),max(x),min(y),max(y)))
        x1 <- pp_s[["x"]]
        y1 <- pp_s[["y"]]
        df_xy <- in_cell_to_c(x1, y1, u, t)
        set_B <- c(set_B, df_xy)
    }
    pned_A <- PNED_c(set_A)
    pned_B <- PNED_c(set_B)
    L_outcomes[j] <- ks.test(pned_A, pned_B)$p.value < 0.05
}
```

```{r}
cat("HPP - Rate at which H0 was rejected when it was true:", mean(L_outcomes))
```

## Thomas
```{r}
Thetas <- seq(0,2*pi,length.out=1000)
LT_outcomes <- rep(0,100)
for (j in 1:100){
    set_A <- c()
    for (p in 1:20){
        u <- rev(sort(runif(3)))
        t <- runif(3)
        r <- r_mag(Thetas, u, t)
        x <- r * cos(Thetas)
        y <- r * sin(Thetas)
        cell_area <- (0.5*sum(u^2)+9)*pi
        pp_s <- rThomas(7/(cell_area), sqrt(cell_area)*0.05, 3, 
                        win = owin(c(min(x),max(x)),c(min(y),max(y))))
        x1 <- pp_s[["x"]]
        y1 <- pp_s[["y"]]
        df_xy <- in_cell_to_c(x1, y1, u, t)
        set_A <- c(set_A, df_xy)
    }
    set_B <- c()
    for (p in 1:20){
        u <- rev(sort(runif(3)))
        t <- runif(3)
        r <- r_mag(Thetas, u, t)
        x <- r * cos(Thetas)
        y <- r * sin(Thetas)
        cell_area <- (0.5*sum(u^2)+9)*pi
        pp_s <- rThomas(7/(cell_area), sqrt(cell_area)*0.05, 3, 
                        win = owin(c(min(x),max(x)),c(min(y),max(y))))
        x1 <- pp_s[["x"]]
        y1 <- pp_s[["y"]]
        df_xy <- in_cell_to_c(x1, y1, u, t)
        set_B <- c(set_B, df_xy)
    }
    pned_A <- PNED_c(set_A)
    pned_B <- PNED_c(set_B)
    LT_outcomes[j] <- ks.test(pned_A, pned_B)$p.value < 0.05
}
```


```{r}
cat("Thomas - Rate at which H0 was rejected when it was true:", mean(LT_outcomes))
```



# Power of Test
Simulation Study to find power of test for different sigmas.
```{r}
Thetas <- seq(0,2*pi,length.out=1000)
sigmas <- c(0.01, 0.05, 0.075, 0.1, 0.115, 0.125, 0.15, 0.175, 0.2, 0.25, 0.3, 
            0.4, 0.5)
powers2 <- rep(0,13)
for (i in 1:13){
    outcomes <- rep(0,50)
    for (j in 1:50){
        set_A <- c()
        for (p in 1:20){
            u <- rev(sort(runif(3)))
            t <- runif(3)
            r <- r_mag(Thetas, u, t)
            x <- r * cos(Thetas)
            y <- r * sin(Thetas)
            cell_area <- (0.5*sum(u^2)+9)*pi
            pp_s <- rpoispp(50/(cell_area), win=c(min(x),max(x),min(y),max(y)))
            x1 <- pp_s[["x"]]
            y1 <- pp_s[["y"]]
            df_xy <- in_cell_to_c(x1, y1, u, t)
            set_A <- c(set_A, df_xy)
        }
        set_B <- c()
        for (p in 1:20){
            u <- rev(sort(runif(3)))
            t <- runif(3)
            r <- r_mag(Thetas, u, t)
            x <- r * cos(Thetas)
            y <- r * sin(Thetas)
            cell_area <- (0.5*sum(u^2)+9)*pi
            pp_s <- rThomas(10/(cell_area), sqrt(cell_area)*sigmas[i], 5, 
                            win = owin(c(min(x),max(x)),c(min(y),max(y))))
            x1 <- pp_s[["x"]]
            y1 <- pp_s[["y"]]
            df_xy <- in_cell_to_c(x1, y1, u, t)
            set_B <- c(set_B, df_xy)
        }
        pned_A <- PNED_c(set_A)
        pned_B <- PNED_c(set_B)
        outcomes[j] <- ks.test(pned_A, pned_B)$p.value < 0.05
    }
    powers2[i] <- mean(outcomes)
}
```


```{r}
t(data.frame(sigmas=sigmas, powers=powers2))
```


```{r}
# Figure 3.9
plot(sigmas, powers2, pch=16, main=expression("Power of the Test as a function of" ~sigma~ "for Irregular Cells"), xlab = expression(~sigma), ylab = "power")
lines(sigmas, powers2)
```

